---
title: "Stat Computing HW2"
author: "B04303061 丁立恆"
date: "2021/3/27"
output: word_document
---

```{r, include=FALSE}
library(ggplot2)
```

Q1.
In this question, I proposed $f_1 = \dfrac{1}{\sqrt{2\pi}} e^{-\dfrac{(x-1)^2}{2}}$ and $f_2 = \dfrac{1}{\sqrt{2\pi}} e^{-\dfrac{x^2}{2}}$ as envelopes.  

First, we visualize those candidates and our target function by density plot.
As the result, I think the variance of applying f1 will be smaller because its curve is more similar to g(x) than f2.
```{r}
target = data.frame(x = seq(1, 20, 0.1),
                    density = sapply(seq(1, 20, 0.1), 
                                     function(x) x^2/sqrt(2*pi)*exp(-x^2/2)))

ggplot(target)+
  geom_line(aes(x = x, y = density),color="blue")+
  stat_function(fun=dnorm, args=list(mean=1, sd=1), color="red")+
  stat_function(fun=dnorm, args=list(mean=0, sd=1), color="black")
```

Because $\int_{1}^{\infty} \frac{x^2}{\sqrt{2\pi}} e^{-\frac{x^2}{2}} dx$ is $\frac{1}{2} - \int_{0}^{1} \frac{x^2}{\sqrt{2\pi}} e^{-\frac{x^2}{2}} dx$, we 
calculate $\frac{1}{2} - \int_{0}^{1} \frac{g(x)}{f(x)} f(x) dx$ to get $\int_{1}^{\infty}g(x)dx$<br>

In this case, $\frac{g(x)}{f_1(x)} = x^2 e^\frac{-2x+1}{2}$, $\frac{g(x)}{f_2(x)} = x^2$
```{r}
set.seed(20210327)
draw <- 10^5

# For f1, N(1,1)
gf1 <- function(x) x^2 * exp(- (2*x+1) / 2)
norm1 <- sapply(1:draw, function(x){
  while(TRUE){
    temp <- rnorm(1,1,1)
    if(temp >=0 & temp <= 1)
      return(temp)
  }
})
integral1 <- sapply(norm1, gf1)

# For f2, N(0,1)
gf2 <- function(x) x^2
norm2 <- sapply(1:draw, function(x){
  while(TRUE){
    temp <- rnorm(1,0,1)
    if(temp >=0 & temp <= 1)
      return(temp)
  }
})
integral2 <- sapply(norm2, gf2)

se <- round(c(sd(integral1), sd(integral2)), 4)
names(se) <- c('se for f1', 'se for f2')
se
```

Estimate $\int_1^{\infty} g(x)dx$ by important sampling
```{r}
0.5 - mean(integral1)
```

Q2. 
First, we calculate the optimal constant c in control variate. Here f is the integral we want to estimate, and g is the control variate.
```{r}
set.seed(20210327)
draw <- 10^5
uniform_var <- runif(draw, min=0, max=10)
f <- function(x) x / (2 + x^0.25)
g <- function(x) x
f_x <- sapply(uniform_var, f)
g_x <- sapply(uniform_var, g)
c <- - cov(f_x, g_x) / var(g_x)
```

Second, we estimate the integral by control variate
```{r}
# 5 is the theoretical expected value for E(g(x))
# (10 - 0) is because the integrate area is 0 to 10
control_integral <- mean(f_x + c*(g_x - 5)) * (10 - 0)
control_integral
```

Third, we simulate the integral by Monte Carlo simulation.
```{r}
set.seed(20210327)
uniform_var <- runif(draw, min=0, max=1)
monte_f <- function(x) 100*x / (2 + (10*x)^0.25)
monte_f_x <- sapply(uniform_var, monte_f)
mean(monte_f_x)
```

Last, we calculate the variance reduction by using control variate
The theoretical reduction is $-Cov(f(x),g(x))^2/Var(g(x)))$
```{r}
monte_var <- var(monte_f_x)
control_var <- var((f_x + c*(g_x - 5))*10)
(monte_var - control_var) / monte_var
```